### 2017-03-01 docprep/appdx_a.tex
Updated the literature review section

%%
%% This is file `chapmin.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ths.dtx  (with options: `chapmin')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from chapmin.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file ths.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)

 %% ... sample chapter ...

This appendix covers the math for Legendre polynomials and spherical harmonics

\section{Legendre Polynomials}

They exist.

\subsection{Spherical Harmonics}

They also exist.

The real harmonics. The even/odd harmonics.

The addition theorem.

\endinput
%%
%% End of file `chapmin.tex'.