### 2017-03-01 docprep/chapbib.tex

%%
%% This is file `chapbib.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ths.dtx  (with options: `chapmin,addbib')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from chapbib.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file ths.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)

 %% ... sample chapter ...

This section is the literature review. I do these things.

\section{Introduction}
Many different things are required.

\section{CT Dose Estimation}
How is this done now?

\section{CT Phantom Generation}
The CT phantom is populated with CT numbers, also known as Hounsfield numbers. These values are related to the attenuation coefficient of the material represented by the voxel.

Many authors have provided correlations that map the CT number to both a material and a density.

First ones.
Saw \citep{ref:sawc} did stuff. And \citep{ref:plessisf}. Then \citet{ref:schneideru} said stuff. And \citet{ref:kimh} did stuff with GEANT4.

Another Schneider did some stuff too \citep{ref:schneiderw}. And Vanderstraeten has a long name \citep{ref:vanderstraetenb}.

The new guy provides 19 materials. The density is measured.

Two phantom datasets are used in this work.

\section{Discrete Ordinates}
Discrete ordinates are old. The book is \citet{ref:lewise}.

The Exnihilo thing by \citet{ref:evanst}. And DeHart wrote in \citep{ref:dehartm}

The Ibrahim did stuff \citep{ref:ibrahima} and Lee did stuff too \citep{ref:leeb}.

With regards to angles, stuff has been done \citep{ref:ahrensc}

Attila had a 1st col src added by \citep{ref:wareingt}. Some rays are better than others \citep{ref:mathewsk}.

Dort and Tort \citep{ref:rhoadesw}.

Denovo.

\section{Raytracing}

Raytracing is important too. Woo came up with on \citep{ref:wooa} which was improved by \citet{ref:liuy} and \citet{ref:hel}.

\section{GPU Acceleration}
GPUs are cool.

CUDA is cool.

\endinput
%%
%% End of file `chapbib.tex'.