### 2017-03-01 docprep/norris_phd_dissertation.tex

 % -*- mode: LaTeX -*-
 % ... preamble ...
\documentclass[times,12pt,titlepage]{mstthesis}
\doublespacing
\usepackage{threeparttable}
\usepackage[final]{graphicx}
\usepackage{psfrag}
\usepackage{listings}
\usepackage{minted}
\usepackage[noprefix]{nomencl}
\makenomenclature
\renewcommand{\nomgroup}[1]{
   \ifthenelse{\equal{#1}{A}}{\medskip \item \textbf{Roman}}{}%
   \ifthenelse{\equal{#1}{G}}{\medskip \item \textbf{Greek}}{}%
   \ifthenelse{\equal{#1}{S}}{\medskip \item \textbf{Subscripts}}{}%
}
\usepackage[nopar]{lipsum} % ... provides dummy text ...

 % ... end preamble ...

\begin{document}

 % ... specify: ms or phd ...

\begin{ThesisTitlePage}{phd}

 % ... title page info ...

\author{\MakeUppercase{Edward T. Norris}}

\thesistitle{\MakeUppercase{Discrete Ordinates CT Organ Dose Simulator (DOCTORS)}}

\department{Nuclear Engineering}

 % ... thesis committee ...

\ThesisAdviser{Xin Liu}

 % If you have a co-advisor, enter the name in the
 % curly braces below and uncomment.
 % Otherwise, leave commented out.
 %\cothadviser{} % If you have 2 thesis advisers

\memberone{Ayodeji Alajo}
\membertwo{Hyoung Koo Lee}
\memberthree{Gary Mueller}
\memberfour{Fikret Ercal}
%\memberfive{Prandtl}

 % ... Graduation date - NOT your submission date ...
\graddate{2017}

\end{ThesisTitlePage}
 % ... copyright page - true|false ...

\copyrightyear{2017}
\ThesisCopyrightPage{true}

 % ... front matter - thesis abstract ...

\begin{ThesisAbstract}
This is the text for the abstract. Here, I explain what I did and how cool I am for doing it. Because I am awesome.
\end{ThesisAbstract}

 % ... front matter - thesis acknowledgements ...

\begin{ThesisAcknowledgment}
I would like to thank these people. They are all so super cool. So much coolness. So much so that this is now over a line long.

And these funding agencies.
\end{ThesisAcknowledgment}
\begin{ThesisFrontMatter}
\tableofcontents
\listoffigures
\listoftables
\listofsymbols
\end{ThesisFrontMatter}

\begin{ThesisBody}

 % ... introduction chapter ...
\ThesisBodyChapter{Introduction}\input{chapintro}
 % ... chapter ...
\ThesisBodyChapter{Literature Review}\input{chapbib}
 % ... chapter ...
\ThesisBodyChapter{Methods}\input{chapmeth}

\ThesisBodyChapter{Implementation}\input{chapimpl}
 % ... conclusions chapter ...
\ThesisBodyChapter{Conclusions}\input{chapconc}

\end{ThesisBody}
 % ... appendix - specify number of appendix chapters ...

\begin{ThesisAppendix}{two}
\ThesisAppendixChapter{Legendre Polynomials and Spherical Harmonics}\input{appdx_a}
\ThesisAppendixChapter{Two}\input{appdx_b}
\end{ThesisAppendix}

 % ... references - comma separated list of bib files ...

\begin{ThesisBibliography}{BIBLIOGRAPHY}
\bibliographystyle{plainnat}
\bibliography{refs}
\end{ThesisBibliography}

 % ... vita ...

\begin{Vita}
Edward Norris graduated from Missouri University of Science and Technology in 2013 with a B.S. in both nuclear engineering and computer science.

He has worked...

He is a member of...
\end{Vita}

\end{document}