### 2017-05-05 docprep/chapbib.tex
Added a new chapter - results.

%%
%% This is file `chapbib.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ths.dtx  (with options: `chapmin,addbib')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from chapbib.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file ths.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)

 %% ... sample chapter ...

This chapter summarizes the literature reviewed in preparation of this work. The literature review is split into multiple sections, each highlighting relevant works pertaining to a particular component or section of this work. The first sections cover introductory work and steps necessary for preprocessing data. Section~\ref{sec:discordlit} covers historical and recent work in discrete ordinate methodology, the methodology used by DOCTORS. The remaining sections cover GPU hardware and other implementation facets.

\section{CT Dose Estimation}
The dose a patient receives form both diagnostic CT procedures and radiation therapy is of great concern to the medical community [CITE].

Monte Carlo codes have been used in the past to compute flux-to-dose conversion factors for a reference person [CITE ICRP].

\section{CT Phantom Generation}
The CT phantom is populated with CT numbers, also known as Hounsfield numbers. These values are related to the attenuation coefficient of the material represented by the voxel.

Many authors have provided correlations that map the CT number to both a material and a density.

First ones.
Saw \citep{ref:sawc} did stuff. And \citep{ref:plessisf}. Then \citet{ref:schneideru} said stuff. And \citet{ref:kimh} did stuff with GEANT4.

Another Schneider did some stuff too \citep{ref:schneiderw}. And Vanderstraeten has a long name \citep{ref:vanderstraetenb}.

The new guy provides 19 materials. The density is measured.

Two phantom datasets are used in this work [CITE].

\section{Discrete Ordinates}\label{sec:discordlit}
Discrete ordinates are old. The book is \citet{ref:lewise}.

The Exnihilo thing by \citet{ref:evanst}. And DeHart wrote in \citep{ref:dehartm}

The Ibrahim did stuff \citep{ref:ibrahima} and Lee did stuff too \citep{ref:leeb}.

With regards to angles, stuff has been done \citep{ref:ahrensc}

Attila had a 1st col src added by \citep{ref:wareingt}. Some rays are better than others \citep{ref:mathewsk}.

Dort and Tort \citep{ref:rhoadesw}.

Denovo.

\section{Raytracing}

Raytracing is important too. Woo came up with on \citep{ref:wooa} which was improved by \citet{ref:liuy} and \citet{ref:hel}.

\section{GPU Acceleration}
GPUs are cool.

CUDA is cool.

\endinput
%%
%% End of file `chapbib.tex'.