### 2017-05-12 docprep/norris_phd_dissertation.tex

 % -*- mode: LaTeX -*-
 % ... preamble ...
\documentclass[times,12pt,titlepage]{mstthesis}
\doublespacing
\usepackage{threeparttable}
\usepackage[final]{graphicx}
\usepackage{psfrag}
\usepackage{listings}
\usepackage{minted}
\usepackage[noprefix]{nomencl}
\usepackage{subcaption}
\usepackage{xr}
\makenomenclature
\renewcommand{\nomgroup}[1]{
   \ifthenelse{\equal{#1}{A}}{\medskip \item \textbf{Roman}}{}%
   \ifthenelse{\equal{#1}{G}}{\medskip \item \textbf{Greek}}{}%
   \ifthenelse{\equal{#1}{S}}{\medskip \item \textbf{Subscripts}}{}%
}
\usepackage[nopar]{lipsum} % ... provides dummy text ...

 % ... end preamble ...

\begin{document}

 % ... specify: ms or phd ...

\begin{ThesisTitlePage}{phd}

 % ... title page info ...

\author{\MakeUppercase{Edward T. Norris}}

\thesistitle{\MakeUppercase{Discrete Ordinates CT Organ Dose Simulator (DOCTORS)}}

\department{Nuclear Engineering}

 % ... thesis committee ...

\ThesisAdviser{Xin Liu}

 % If you have a co-advisor, enter the name in the
 % curly braces below and uncomment.
 % Otherwise, leave commented out.
 %\cothadviser{} % If you have 2 thesis advisers

\memberone{Ayodeji Alajo}
\membertwo{Hyoung Koo Lee}
\memberthree{Gary Mueller}
\memberfour{Fikret Ercal}
%\memberfive{Prandtl}

 % ... Graduation date - NOT your submission date ...
\graddate{2017}

\end{ThesisTitlePage}
 % ... copyright page - true|false ...

\copyrightyear{2017}
\ThesisCopyrightPage{true}

 % ... front matter - thesis abstract ...

\begin{ThesisAbstract}
This is the text for the abstract. Here, I explain what I did and how cool I am for doing it. Because I am awesome.
\end{ThesisAbstract}

 % ... front matter - thesis acknowledgements ...

\begin{ThesisAcknowledgment}
I would like to thank my adviser, Dr. Liu and all of the other graduate students for their continued support through my graduate career. I also appreciate the time spent by my committee members helping me identify problems and provide insight to solutions.

I would also like to acknowledge the NRC for providing funding for my work. My work is supported by NRC grant NRC-HQ-13-G-38-0026.
\end{ThesisAcknowledgment}

\begin{ThesisFrontMatter}
\tableofcontents
\listoffigures
\listoftables
\listofsymbols
\end{ThesisFrontMatter}

\begin{ThesisBody}

 % ... introduction chapter ...
\ThesisBodyChapter{Introduction}\input{chapintro}
 % ... chapter ...
\ThesisBodyChapter{Literature Review}\input{chapbib}
 % ... chapter ...
\ThesisBodyChapter{Methods}\input{chapmeth}

\ThesisBodyChapter{Implementation}\input{chapimpl}
 % ... conclusions chapter ...

\ThesisBodyChapter{Results}\input{chapresults}

\ThesisBodyChapter{Conclusions}\input{chapconc}

\end{ThesisBody}
 % ... appendix - specify number of appendix chapters ...

\begin{ThesisAppendix}{two}
\ThesisAppendixChapter{Legendre Polynomials and Spherical Harmonics}\input{appdx_a}

\ThesisAppendixChapter{Code Listings}\input{appdx_b}

\ThesisAppendixChapter{Benchmark Results}\input{appdx_c}

\end{ThesisAppendix}

 % ... references - comma separated list of bib files ...

\begin{ThesisBibliography}{BIBLIOGRAPHY}
\bibliographystyle{plainnat}
\bibliography{refs}
\end{ThesisBibliography}

 % ... vita ...

\begin{Vita}
Edward Norris graduated from Missouri University of Science and Technology in 2013 with a B.S. in both nuclear engineering and computer science, he went on to pursue a M.S. in computer science and a PhD in nuclear engineering.

His PhD work was funded by a fellowship through the NRC.

Previously he has participated in four summer internships with Sandia National Laboratories. While there, he served in various capacities ranging from cyber defense to radiation detection and neutron scatter simulation.

Edward was awarded SOME AWARDS.

Edward was also an active member of the Nuclear Science Design Team at Missouri University of Science and Technology. While on the team he solved radiation transport problems to facilitate the dosimetry of an inertial electrostatic confintement fusion device being constructed.

\end{Vita}

\end{document}